### 2020-03-23 sections/discussion.tex
Includes new results (deprojection from GZB fitting, updated hyperparameters on clustering). Greatly expands on intro section

The statistical approach implemented in this paper allows a more thorough examination of pitch angle than simpler methods of logarithmic spiral fitting, and better accounts for inter-arm pitch angle variation than fourier analysis \comment{This is a strong statement, how can we quantify it?}.

Our method and results do not completely reject spiral winding of the form described by \citet{2019arXiv190910291P}, however this result is highly sensitive to the boundaries used; using limits of $10.0^\circ < \phi < 40.0^\circ$ results in us rejecting the null hypothesis at the 1\% level for every MCMC draw. As no physical justification was provided for the limits present in \citet{2019arXiv190910291P}, this is a far from ideal test of arm winding.

@@

This work is primarily limited by the sample used; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative. A larger and more complete sample is needed to be completely confident in our results. However, we believe that the methodlogy proposed here is a scalable solution to the problems facing investigation of spiral morphology.

>>>>

The statistical approach implemented in this paper allows a more thorough examination of pitch angle than simpler methods of logarithmic spiral fitting, and better accounts for inter-arm pitch angle variation than fourier analysis, which assumes all arms in a given mode have the same pitch angle.

Our method and results do not completely reject spiral winding of the form described by \citet{2019arXiv190910291P}, however this result is highly sensitive to the boundaries used; using limits of $10.0^\circ < \phi < 40.0^\circ$ results in us rejecting the null hypothesis at the 1\% level for 99\% of the MCMC draws. As no physical justification was provided for the limits present in \citet{2019arXiv190910291P}, this is a far from ideal test of arm winding.

@@

As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative. However, we believe that the methodlogy proposed here is a scalable, robust solution to the problems facing investigation of spiral morphology.

### 2020-03-23 sections/introduction.tex
Includes new results (deprojection from GZB fitting, updated hyperparameters on clustering). Greatly expands on intro section

% BEGIN HERE
Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated. Spirals are as diverse as the theories proposed to govern their evolution; from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral. These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}). Our current understanding of the mechanisms which drive spiral growth and evolution suggest that each of the different forms of spiral galaxy may be triggered primarily by different processes. Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as suggested by Manifold theory, \citealt{2011hsa6.conf..548R}), or be obeying quasi-stationary density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}. Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}. It is recognised that methods of spiral formation are not mutually exclusive. There is also debate as to the evolution of spiral structure: whether spirals maintain their shape over many rotational periods (assumed by QSDW theory, \citealt{1964ApJ...140..646L}), or ``wind up'' with the differential rotation of the disc (as found by \citealt{2019MNRAS.487.1808M}).

@@

@@

Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), fourier analysis (i.e. \textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e. SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers \citep{2017MNRAS.472.2263H}. One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\comment{Lingard et al, in prep}). These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization. This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}. \comment{Lingard et al, in prep} propose a solution to this problem through the use of citizen science to provide the starting points for computational fitting.

@@

The most commonly used measurement of uncertainty of length-weighted pitch angles is the \comment{unweighted?} sample variance between the arm segments identified.

@@

Fourier analysis in one-dimension (as performed by \comment{...}) and two-dimensions (\comment{...}) is another widely used method of computationally obtaining galaxy pitch angles. Two-dimensional fourier methods generally decompose a deprojected image of a galaxy into a superpositions of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D}, and reports the pitch angle with the highest amplitude as the galaxy's pitch angle. It is uncertain how the large observed variation between pitch angles of individual arms impacts this measurement.

This paper makes use of the photometric models obtained through the \textit{Galaxy Builder} citizen science project for the 196 spiral galaxies present in \comment{Lingard et al, in prep}. In this paper we focus on the use of measured spiral tightness (quantified using pitch angle, \citealt{1987gady.book.....B}) as a probe into the dynamical mechanisms governing a spiral galaxy's evolution. We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters producted by \textit{Galaxy Builder}.

Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \citealt{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and concludes using a marginalized Anderson-Darling test that we \comment{cannot unilaterally reject winding of this form at the 1\% level}. Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size implied by the Hubble sequence and reported by \comment{citation}, and pitch angle and bar strength implied by Manifold theory, and find \comment{no significant correlation}.

>>>>

Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated. Spirals are as diverse as the theories proposed to govern their evolution; from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral. These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}). Our current understanding of the mechanisms which drive spiral growth and evolution suggest that each of the different forms of spiral galaxy may be triggered primarily by different processes. Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as seen in gas simulations, \citealt{1976ApJ...209...53S}, \citealt{2008A&A...489..115R}, and suggested for stars by Manifold theory, \citealt{2006A&A...453...39R}, \citealt{2009MNRAS.394...67A}, \citealt{2009MNRAS.400.1706A}), or be obeying (quasi-stationary) density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}. Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}. It is recognised that methods of spiral formation are not mutually exclusive. There is also debate as to the evolution of spiral structure: whether spirals maintain their shape over many rotational periods (assumed by QSDW theory), or ``wind up'' with the differential rotation of the disc (as found by \citealt{2003MNRAS.344..358B}, \citealt{2019MNRAS.487.1808M}).

@@

\subsection{Secular evolution of spiral arms}

\comment{I'm worried that a lot of this section repeats what was said earlier}

One of the foundational assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanging for many rotational periods \citep{1964ApJ...140..646L}. The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most galaxies show spiral structure, indicating that spirals exist for a long time or are continually rebuilt.

Many simulations demonstrate that spirals do not maintain a constant pitch angle, and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}. Recent research suggests that spirals are dynamical in nature, and continually dissapate and re-form \citep{2014PASA...31...35D}. These spirals can be maintained through the same mechanisms that drive QSDW spirals (i.e. WASER, \citealt{1976ApJ...205..363M}, swing amplification, \citealt{1965MNRAS.130..125G}), but do not require the idealistic disc conditions required for the formation and maintenance of a stationary wave. The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disk, with the density of the arm peaking at some critical pitch angle and then dissapating to be reformed.

In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissapation. \citet{2019arXiv190910291P} proposes a simple test of spiral arm winding, assuming the cotangent of the pitch angle evolves linearly with time. Finding compelling evidence of this relationship against QSDW theory, in favour of the dynamic spirals now favoured by many simulations.

Spiral evolution also appears to be influenced by the presence and strength of a bar; in barred grand-design spirals the arms often appear to start from the ends of the bar. Simulations of gas in barred galaxies often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}. Manifold theory is one attempt to determine the orbits of stars in bar-driven spiral arms: it proposes that stars in the vicinity of the unstable Lagrangian points at either end of the bar tend to escape along predictable orbits, governed by invariant manifolds. One of the primary factors influencing the shape of this invariant manifold is the relative strength of the non-axisymmetric forcing caused by the bar, with stronger bars resulting in spirals with larger pitch angles.

Many other systems contribute to spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2015PhDT........14D}, \citealt{2017MNRAS.471.2187D}), via a series of correlations. Stronger bulges and more more massive central black holes are both linked to more tightly wound spiral arms, albeit with large amounts of scatter.

@@

Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), fourier analysis (i.e. \textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e. SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers (\citealt{2017MNRAS.472.2263H}, \citealt{2020MNRAS.493.3854H}). One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\Lingard). These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization. This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}. \Lingard propose a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.

@@

The most commonly used measurement of uncertainty of length-weighted pitch angles is the unweighted sample variance between the arm segments identified.

@@

Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}) is another widely used method of computationally obtaining galaxy pitch angles. Two-dimensional fourier methods generally decompose a deprojected image of a galaxy into a superpositions of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D}, and reports the pitch angle with the highest amplitude as the galaxy's pitch angle. It is uncertain how the large observed variation between pitch angles of individual arms impacts this measurement.

This paper makes use of the classification data and fitted photometric models obtained through the \textit{Galaxy Builder} citizen science project for the 196 spiral galaxies present in \Lingard. In this paper we focus on the use of measured spiral tightness (quantified using pitch angle, \citealt{1987gady.book.....B}) as a probe into the dynamical mechanisms governing a spiral galaxy's evolution. We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters producted by \textit{Galaxy Builder}.

Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \cite{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and concludes using a marginalized Anderson-Darling test that we cannot unilaterally reject winding of this form at the 1\% level. Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory, and find no significant correlation.

### 2020-03-23 sections/method.tex
Includes new results (deprojection from GZB fitting, updated hyperparameters on clustering). Greatly expands on intro section

% !TEX root = ../main.tex
\subsection{The Galaxy Sample}
The galaxies selected analysed in this paper are the 198 galaxies from \comment{Lingard et al, in prep.}. We combine classifications of galaxies which were repeated in the \textit{validation subset} with the original classifications. Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \comment{Lingard et al, in prep.}. We remove any galaxies for which no arms were identified. This results in a hierarchiacal data structure of 109 galaxies, 250 spiral arms and 307,861 points.

Spiral arm points are deprojected to a face-on orientation using the position angle \textsc{PETRO\_PHI90} and axis ratio \textsc{PETRO\_BA90} keywords from the NASA-Sloan Atlas \comment{citation}. We scale the radii to have unit variance (which does not impact log spiral pitch angle due to its scale invariance).

\comment{Should I elaborate more on the sample?}

@@

\begin{figure}
  \includegraphics[width=8cm]{plots/plots_n109d1000t500/model.pdf}
  \caption{The model used for galaxy pitch angle measurement for the sample.}
  \label{fig:pymc3-model}
\end{figure}

>>>>

% !TEX root = ../main.tex
\subsection{The Galaxy Sample}
The galaxies selected analysed in this paper are the 198 galaxies from \Lingard. These are a subset of the \textit{stellar mass-complete sample} in \citet{2017MNRAS.472.2263H}, a sample of low-redshift face-on spiral galaxies selected using data from the NASA-Sloan Atlas \citep{2011AJ....142...31B} and Galaxy Zoo 2 \citep{Willett2013:1308.3496v2}.

We combine classifications of galaxies which were repeated in the \textit{validation subset} with the original classifications. Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \Lingard. We remove any galaxies for which fewer than two spiral arms were identified \comment{why exclude one-armed?}. This results in a hierarchical data structure of 91 galaxies, 211 spiral arms and 215,678 points.

Spiral arm points are deprojected to a face-on orientation using the disk inclination and position angle obtained through photometric model fitting performed in \Lingard. We scale the deprojected point radii to have unit variance, to aid chain convergence.

@@

% \begin{figure}
%   \includegraphics[width=8cm]{plots/plots_n109d1000t500/model.pdf}
%   \caption{The model used for galaxy pitch angle measurement for the sample.}
%   \label{fig:pymc3-model}
% \end{figure}

### 2020-03-23 sections/results.tex
Includes new results (deprojection from GZB fitting, updated hyperparameters on clustering). Greatly expands on intro section

% !TEX root = ../main.tex
\subsection{Constraints on Galaxy Pitch angle}
Our hierarchical model can predict individual arm pitch angle with a hich degree of certainty, however it returns a large spread of potential values for the pitch angles of arms

@@

We perform the marginalized Anderson-Darling test for a potential source distribution uniform in $\cot\phi$ between the limits present in \citet{2019arXiv190910291P} ($1.00 < \cot\phi < 4.75$, or roughly $11.9^\circ < \phi < 45.0^\circ$). The resulting distribution of Anderson-Darling statistics can be seen in Figure \ref{fig:ad-cot-test}. We observe that we reject the null hypothesis at the 1\% level for only 86\% of the possible realizations of galaxy pitch angle, therefore with our sample and methodology we cannot unilaterally reject winding of the kind described by \citet{2019arXiv190910291P}.

@@

@@

We separate our sample into ``disc-dominated galaxies'' and ``obvious bulge galaxies'' using the debiased fractions from Galaxy Zoo 2 following \citet{2017MNRAS.469.3363K}, defining \textit{no bulge + just noticeable $>$ obvious + dominant} for the former and the converse for the latter. A marginalized two-sample Anderson-Darling test \citep{doi:10.1080/01621459.1987.10478517} does not find strong evidence that the samples were drawn from different distributions; we reject the null hypothesis at the 1\% level for only 8\% of the samples.

@@

Separating the sample based off of $\mathrm{\textit{pbar}} > 0.5$, and restricting to galaxies with more than 10 classifications for \textit{pbar} (as performed by \citealt{2011MNRAS.411.2026M} and \citealt{2017MNRAS.469.3363K}) and performing a marginalized two-sample Anderson-Darling test does not find that the samples were drawn from different distributions. \comment{Talk about number of galaxies in each sub-sample?}

>>>>

% !TEX root = ../main.tex
\subsection{Constraints on Galaxy Pitch angle}
Our hierarchical model identifies arm pitch angle with a hich degree of certainty (uncertainty less than 1.6\degree for 95\% of arms, assuming no error on disc inclination and position angle), however it returns a large spread of potential values for the pitch angles of galaxies, primarily caused by only measuring pitch angles of a small number of arms per galaxy. For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of 7.8\degree, which decreases to 6.7\degree and 5.8\degree for galaxies with three and four arms respectively.

Our measure of inter-arm variability of pitch angle, $\sigma_\mathrm{gal} = 11.02 \pm 0.95$, confirming

@@

We perform the marginalized Anderson-Darling test for a potential source distribution uniform in $\cot\phi$ between the limits present in \citet{2019arXiv190910291P} ($1.00 < \cot\phi < 4.75$, or roughly $11.9^\circ < \phi < 45.0^\circ$). The resulting distribution of Anderson-Darling statistics can be seen in Figure \ref{fig:ad-cot-test}. We observe that we reject the null hypothesis at the 1\% level for 69\% of the possible realizations of galaxy pitch angle. While, with our sample and methodology, we cannot unilaterally reject winding of the kind described by \citet{2019arXiv190910291P}, it suggests this model is not found in our data.

@@

Not seeing uniformity in $\cot\phi$ is not evidence against spiral winding, nor is it evidence for static spiral structures. Our results simply suggest that a more comprehensive test is needed for observational data of this kind.

@@

We separate our sample into ``disc-dominated galaxies'' and ``obvious bulge galaxies'' using the debiased fractions from Galaxy Zoo 2 following \citet{2017MNRAS.469.3363K}, defining \textit{no bulge + just noticeable $>$ obvious + dominant} for the former and the converse for the latter. A marginalized two-sample Anderson-Darling test \citep{doi:10.1080/01621459.1987.10478517} does not find any evidence that the samples were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples. The distribution of Anderson-Darling test statistics is shown in the upper panel of Figure \ref{fig:ad-morphology-test}.

@@

Separating the sample based off of $\mathrm{\textit{pbar}} > 0.5$, and restricting to galaxies with more than 10 classifications for \textit{pbar} (as performed by \citealt{2011MNRAS.411.2026M} and \citealt{2017MNRAS.469.3363K}) and performing a marginalized two-sample Anderson-Darling test does not find that the samples were drawn from different distributions (we reject the null hypothesis at the 1\% level for only 1\% of samples). The distribution of Anderson-Darling test statistics is shown in the lower panel of Figure \ref{fig:ad-morphology-test}.

We do not account for variation in bulge size in this work, however predictions from Manifold theory should not be affected as bulges do not provide a non-axisymmetric focing.
